Review: Redefining Domestication: Adaptive Introgression during Crop Expansion

### 2017-06-14 paper.tex
re-worked the rice section.

The story of Asian rice domestication is still debated.
A recent investigation pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{Oryza rufipogon}, with later divergence of the two prominent varietal groups, indica and japonica \cite{molina2011molecular}.
On the other hand, both genetic and archaeobotanical evidence point toward independent domestications of japonica and indica in the Yangtze Basin and the Indian Ganges plain respectively \cite{fuller2010consilience}.
An alternative explanation, as described by \cite{Huang2012}, suggests an initial domestication of japonica in China followed by formation of the indica group through hybridization of japonica and local \emph{O. rufipogon} populations in Southern and South-eastern Asia.
As suggested by \cite{vaughan2008evolving}, gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits.
Adaptive introgression may in fact be the signal interpreted as independent domestications.

The wild relatives rufipogon and nivara are both phenotypically diverse and show evidence of local adaptation.
Rufipogon reproduction can be either primarily sexual or vegetative.
Whereas rufipogon is adapted to forested wetland environments, nivara is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure.
Likewise, cultivated rice varieties display phenotypic differences important for fitness.
In particular, two of the cultivated rice deepwater varieties (rayada and ashwina) are said to be selected for the environment along the Ganges river.
Gene flow between these wild and cultivated rices is common, producing nuisance weedy hybrids in when the two are grown in sympatry.
However, during early domestication, introgression may have been even more prevalent because barriers to crop-to-wild introgression may have been less severe and because the inbreeding reproductive system of rice would not have been as firmly established.
Furthermore, the contemporary distribution of wild rice does not capture the range and diversity of wild rice during early domestication and range expansion of rice.
Introgression into indica is expected to be more likely than japonica, due to the higher degree of sympatry between indica and its wild relatives.
The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}.

@@

>>>>

@@

The story of Asian rice (\emph{Oryza sativa}) domestication is still debated.
A recent investigation pointed to a single domestication occurring 8,200-13,500 BP in the Yangtze Basin in China from the wild species \emph{Oryza rufipogon} (\emph{rufipogon} hereafter), with later divergence of the two prominent varietal groups, \emph{japonica} and \emph{indica} \cite{molina2011molecular}.
On the other hand, both genetic and archaeobotanical evidence point toward independent domestications of \emph{japonica} and \emph{indica} in the Yangtze Basin and the Indian Ganges plain, respectively \cite{fuller2010consilience}.
An alternative explanation, as described by \cite{Huang2012}, suggests an initial domestication of \emph{japonica} in China followed by formation of the \emph{indica} group through hybridization of \emph{japonica} and local \emph{rufipogon} populations in Southern and South-eastern Asia.
Adaptive introgression may in fact be the signal interpreted as independent domestications.

The high genetic diversity within domesticated rice is likely due to introgression from wild relatives both within the domestication center(s) and in new environments where rice has dispersed \cite{second1982origin}, and gene flow between domesticated Asian rice and its wild relatives outside of the Yangtze Basin could have imparted locally adaptive traits \cite{vaughan2008evolving}.
The wild relatives \emph{rufipogon} and \emph{nivara} both maintain high phenotypic diversity and exhibit locally-adaptive traits.
\emph{Rufipogon} reproduction can be either primarily sexual or vegetative, and whereas \emph{rufipogon} is adapted to forested wetland environments, \emph{nivara} is adapted to dryer conditions and has life cycle adaptations to survive grazing pressure.
Likewise, domesticated rice varieties display patterns of local adaptation.
For examples, two of the domesticated rice deepwater varieties (\emph{rayada} and \emph{ashwina}) are said to be selected for the environment along the Ganges river, the \emph{japonicas} are split into temperate and tropical subgroups, and the \emph{indicas} are best suited for lowland environments.
Gene flow between these wild and cultivated rices, though asymmetrical, is exceedingly common, producing nuisance weedy hybrids when the two are grown in sympatry.

To date, research into adaptive introgression in the domestication of rice has been insufficient to detect clearly-supported examples.
\cite{zhao2010genomic} used a SNP panel and STRUCTURE analysis to uncover patterns of population structure, admixture, and introgrogression within domesticated rices, and the authors emphasize the importance of similar research that includes wild rice accessions.
There are perhaps some practical reasons why research has not yet been devoted to this inquiry.
As with many other domesticated crops, gene flow between wild and domesticated rices is highly asymmetric (estimates of wild rice admixture in domesticated rice are less than 5 percent \cite{wang2017asian}).
This asymmetry is due in part to the closed floret architecture of the domesticated rice, which hinders outcrossing.
However, during early domestication, introgression may have been more prevalent than at present because barriers to crop-to-wild introgression may have been less severe and because the inbreeding reproductive system of rice would not have been as firmly established \cite{vaughan2008evolving}.
Furthermore, the contemporary distribution of wild rice does not capture the range and diversity of wild rice during early domestication and range expansion of rice.

%Introgression into indica is expected to be more likely than japonica, due to the higher degree of sympatry between indica and its wild relatives.
